---
title: "Disease"
author: "Faris Dzikrur R"
date: "31/07/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readxl)
library(janitor)
library(broom)

library(rjson)
library(bbplot)
library(ggthemes)
library(maps)
library(leaflet)
library(mapdata)
library(plotly)
library(bbplot)

library(tidyverse)

library(ggthemes)
library(scico)

library(mapproj)
library(animation)
library(tweenr)
library(gganimate)
library(ggthemes)
```

Global Mortality:

WHO: Those who concern with public health policy issue

WHAT:  
1. Top cause of death in each country
2. Trend of each disease across the year in every country
3. Correlation between cause of death and life expectancy

WHY:
1. 

WHERE:
1. Global

WHEN:
1990 - 2016

HOW:
1. I use bar plot to know the top cause of death in each country
2. I plot the death rate by using map to know where each of the death cause spread in each country across the years
3. I use geom_point to make a correlation between death rate and life expectancy so we could know which disease have negative correlation with the growth in life expectancy.
```{r}
global_mortality <- read_xlsx("global_mortality (1).xlsx")
life_dat <- read_csv("Global_Mortality_Disease/week14_global_life_expectancy.csv") %>% 
  filter(year == 2015) %>% 
  mutate(country = case_when(
    country == "United States" ~ "United States of America",
    country == "Democratic Republic of Congo" ~ "Democratic Republic of the Congo",
    country == "Congo" ~ "Republic of the Congo",
    country == "Cote d'Ivoire" ~ "Ivory Coast",
    country == "Serbia (including Kosovo)" ~ "Republic of Serbia",
    country == "Tanzania" ~ "United Republic of Tanzania",
    TRUE ~ country
  ))
```

```{r}
tidy_mortality <- global_mortality %>% 
  clean_names %>% 
  gather("cause", "percent_mortality", -country, -country_code, -year) %>% 
  mutate(cause = cause %>% str_replace("_percent", ""))

head(tidy_mortality)
```

```{r}
str(tidy_mortality)
```

```{r}
tidy_mortality <- tidy_mortality %>% 
  mutate_if(is.character, as.factor)

str(tidy_mortality)
```

###Plot 1
```{r}
data_plot_arrange <- tidy_mortality %>% 
  filter(year == 2016, country == "China") %>% 
  group_by(cause, year, country) %>% 
  summarise(rates=mean(percent_mortality)) %>% 
  arrange(desc(rates))

ggplot(data_plot, aes(x= reorder(cause, rates), y=rates, fill=cause)) +
  geom_bar(show.legend = F, stat = "identity") +
  
  coord_flip() +

  theme_classic()+
  
  geom_text(aes(label=paste0(round(rates,1), '%')), position = position_dodge(width = .9), hjust = 'left', vjust = 'center', size = 2.5)
  
  
  
  geom_text(aes(label = paste(round(100 * rates, 2), "%")), hjust = -0.1)+
      scale_y_continuous(labels = scales::percent, limits = c(0, 0.35))+
      scale_fill_manual(values = colors)+
      guides(fill = FALSE)+
      coord_flip()+
      xlab("")+
      ylab("")+
      theme_classic()+
      labs(title = title, subtitle = subtitle, caption = caption) 
     

head(data_plot_arrange$rates,1)





  
  
```

###Plot 2
```{r}
mortality <- global_mortality

names(mortality) <- str_trim(str_remove_all(names(mortality), "[[:punct:]]"))

mortality_SDI <- mortality %>% 
  gather(cause, pct, -c("country", "countrycode", "year")) %>%  
  filter(str_detect(country, "SDI"), cause == "Cancers") %>% 
  select(1,3,4,5) %>% 
  mutate(country == factor(country, levels=c("High SDI", "High-middle SDI", "Middle SDI", "Low-middle SDI", "Low SDI")))
```

```{r}
ggplot(mortality_SDI, aes(year, pct/100, color=country)) + 
  geom_line(size=1.2) +
  scale_color_ptol("Socio-Demographic Index") +
  theme_minimal() +
  scale_y_continuous(labels=scales::percent) +
  scale_x_continuous(breaks = c(1990, 1995, 2000, 2005, 2010, 2015)) +
  labs(x=NULL, y=NULL, title="Percentage of Deaths Attributed to Cancer", subtitle = "Not a sign of decreasing")+
  theme_economist() + scale_colour_economist()

```


### Plot 3
```{r}
world <- map_data("world") %>% 
  rename(country = region)

world_mart <- tidy_mortality %>% 
  inner_join(world, by = "country") %>% 
  filter(year == 2015, cause == "cardiovascular_diseases") %>% 
  mutate(z_score = ave(percent_mortality, cause, FUN = scale))
```

```{r}
map <- ggplot(world_mart, aes(x=long, y=lat, group=group)) +
  geom_polygon(aes(fill=percent_mortality), color = "black") +
  coord_fixed(1.3) +
  scale_fill_gradient(high="red", low="yellow") +
  labs(title="Cardiovascular Disease Rates Across The Globe in 2015")

ggplotly(map)
```

### Plot 4

```{r}
big_dat <- left_join(life_dat, tidy_mortality, by=c("code"="country_code", "year"="year"))

test <- big_dat %>% 
  filter(!is.na(code), cause == "cardiovascular_diseases")

ggplot(test, aes(x = percent_mortality/100, y = life_expectancy)) + 
  geom_point(alpha = .5) + 
  geom_smooth(aes(color = cause)) + 
  scale_color_viridis_d() +
  labs(title = "Percent of Deaths to Life Expectancy Worldwide",
       y = "Life Expectancy", x = "Percent") +
  guides(color = FALSE) 
```


```{r}
mortality_SDI_1 <- mortality %>% 
  gather(cause, pct, -c("country", "countrycode", "year"))

mortality_SDI_1 <- mortality_SDI_1 %>% 
  mutate_if(is.character, as.factor)

p <- mortality_SDI_1 %>% 
  filter(str_detect(country, "SDI"), cause == "Cancers") %>% 
  select(1,3,4,5) %>% 
  mutate(country == factor(country, levels=c("High SDI", "High-middle SDI", "Middle SDI", "Low-middle SDI", "Low SDI")))

check <- mortality_SDI_1 %>% 
  filter(str_detect(country, "SDI")) %>% 
  select(1,3,4,5)

lsm <- check %>% 
  filter(cause == "Cancers")

ggplot(lsm, aes(year, pct/100, color=country)) + 
  geom_line(size=1.2) +
  scale_color_ptol("Socio-Demographic Index") +
  theme_minimal() +
  scale_y_continuous(labels=scales::percent) +
  scale_x_continuous(breaks = c(1990, 1995, 2000, 2005, 2010, 2015)) +
  labs(x=NULL, y=NULL, title="Percentage of Deaths Attributed to Cancer", subtitle = "Not a sign of decreasing")


mortality_final <- tidy_mortality %>% 
  filter(str_detect(country, "SDI")) %>% 
  select(1,3,4,5)

data_output_2 <- mortality_final %>% 
  filter(cause == "major_options_2")
```


```{r}
check$cause


lsm$country

```

```{r}
   major_react_2 <- reactive({
       major_options_2 <- input$major_options_2
      
      data_output_2 <- mortality_final %>% 
         filter(cause == "major_options_2")
       
       return(data_output_2)
   })
   
   output$major_plot_7 <- renderPlot({
       data_input_2 <- major_react_2()
       
       
       plot_major_7 <- ggplot(data_input_2, aes_string(x=year,y=pct/100, color = country)) + 
           geom_line(size=1.2) +
           scale_color_ptol("Socio-Demographic Index") +
           theme_minimal() +
           scale_y_continuous(labels=scales::percent) +
           scale_x_continuous(breaks = c(1990, 1995, 2000, 2005, 2010, 2015)) +
           labs(x=NULL, y=NULL, title="Percentage of Deaths Attributed to Disease") 
       
       ggplot(plot_major_7)
   })
```

```{r}
 menuItem(
            text = "Socio-Demographic Index",
            tabName = "SDI",
            icon = icon("chart-line")
        )



```

```{r}
  tabItem("SDI",
                fluidRow(
                    box(
                        "Filter", br(), "Rates trend based on SDI",
                        selectInput(
                            inputId = "major_options_2",
                            label = "Select Disease:",
                            choices = levels(mortality_final$cause),
                            multiple = FALSE
                        )
                    ),
                    plotOutput(outputId = "major_plot_7")
                    
                )
            
        ),
```

